The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2018-04-23 script/en/chapters/numbertheory.tex
CT1: Updated CT book (d+e): typos, dead links, 2018, 50th M-prime, automated rowno in 2 tables in chap 3, Make.rules also for Windows

\noindent Concretely for the RSA method does this mean:
\vskip - 1em
\begin{enumerate}
  \item It is hard to calculate $\phi(n)$ for big compounds $n$ and
  \item It is hard to calculate the prime factors of big compounds $n$
        (factorization problem).%
\footnote{%
    There is no reason for the concern sometimes mentioned that there are
    not enough primes: Raising the dimension (exponent) of the modul always
    offers enough primes to consider -- this is visualized in
    chapter~\ref{primes:_Appendix_Plotting-Primes-Quantity}
    ``\nameref{primes:_Appendix_Plotting-Primes-Quantity}''.
}
\index{factorization!factorization problem}.
\end{enumerate}
\vskip +1em

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{SecurityRSA}{}
\section[Considerations regarding the security of the RSA algorithm]
{Regarding the security of the RSA algorithm\footnotemark}
    \footnotetext{%
    Major parts of the first part of chapter~\ref{SecurityRSA} follow
    the article ``Vorz\"uge und Grenzen des RSA-Verfahrens'' written by
    F.~Bourseau, D.~Fox and C.~Thiel \cite{Bourseau2002}.}
    \label{SecurityRSA}

@@

%%_BEFORE_MOVING_OUT_INDEX:  \paragraph*{Factorization algorithms\footnote{%
\footnotetext{%
    \index{NT, Learning Tool for Number Theory}%
    \index{educational tool NT}%
    With the educational tool for number theory {\bf NT} you can gather more
    experience with current factorization algorithms
    (see learning unit 5.1-5.5, pages 1-15/15).\\
    NT can be called in CT1\index{CT1} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
    See appendix~\ref{s:appendix-Learn-NT}.\\
    \noindent The quadratic sieve (QS) can be found in CT1 and CT2;
    GNFS, the most modern factorisation method for moduli bigger than 130
    decimal digits, is only part of CT2 (via YAFU and msieve).\\
  \noindent CT2 \index{CT2} has a GeneralFactorizer component
  based on YAFU\index{YAFU}. This is faster than the implemented functions in CT1.
  So CT2 offers the following factoring methods:
  \begin{compactitem}
   \item[-] brute-force with small primes
   \item[-] Fermat
   \item[-] Shanks square forms factorization (squfof)
   \item[-] Pollard rho
   \item[-] Pollard p-1
   \item[-] Williams p+1
   \item[-] Lenstra elliptic curve method (ECM)
   \item[-] self-initializing quadratic sieve (SIQS)
   \item[-] multiple polynomial quadratic sieve (MPQS)
   \item[-] special number field sieve (SNFS)
   \item[-] general number field sieve (GNFS).         
  \end{compactitem}
}
%%_BEFORE_MOVING_OUT_INDEX:  }
\mbox{}
%\nopagebreak

@@

An exhaustive overview about the factoring records
\index{factorization!factoring records} of composed integers using different
methods can be found on the following web pages:
\vspace{-10pt}
\begin{itemize}
\item[]
     \url{http://www.crypto-world.com}\\
     \url{http://en.wikipedia.org/wiki/Integer_factorization_records}\\
     \url{http://en.wikipedia.org/wiki/RSA_Factoring_Challenge}
\end{itemize}
%\footnote{%
%This site was not quite up-to-date end of May 2005: RSA-200 was missing.}

@@

The 'C numbers' originate from the Cunningham project:
\url{http://homes.cerias.purdue.edu/~ssw/cun/}.
These are factors of Mersenne numbers, which have a very special
form. This  makes it an order of magnitude easier to factor them
as moduli of the same length build for RSA.
                          }
with factorization algorithms for composed numbers are
listed in the table~\ref{factorizationrecords}.

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]   \url{https://members.loria.fr/PZimmermann/records/gnfs158}\\
          \url{http://www.crypto-world.com/announcements/c158.txt}
\end{itemize}
\end{minipage}
\vspace{12pt}

@@

On January 18, 2002 researchers at the German University of Bonn\footnote{%
          \url{https://members.loria.fr/PZimmermann/records/rsa160}\\
          \url{https://members.loria.fr/PZimmermann/records/factor.html}\\
          \url{http://www.crypto-world.com/FactorWorld.html}
} 
factorized a 160-digit number into its both prime factors (these are build
with each 80 decimal digits) using the GNFS method (general number field
sieve)\index{general number field sieve (GNFS)}.

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item[]]  \url{http://www.loria.fr/~zimmerma/records/21039-}\\
          \url{http://www.crypto-world.com/announcements/m1039.txt}\\
          \url{http://www.crypto-world.com/FactorAnnouncements.html}
\end{itemize}
\end{minipage}

@@

\noindent There are three variants:
\begin{itemize}
 \item Powerpoint 2007 (for download; dynamical, animated)\footnote{%
    \url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-de.pptx}}
 \item PDF (for download; static, no interaction)\footnote{%
    \url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-de(keine%20Interaktivitaet).pdf}}
 \item Flash (can be started within the browser, requires JavaScript; time-controlled replay)\footnote{%
\url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-Flash-de/player.html}}
\end{itemize}

@@

   \item Factorization records and factoring challenges
	 \index{factorization!factoring records}
	 \index{factorization!factoring challenges}\\
         \url{http://www.crypto-world.com/},~~ last update 2013\\
         \url{http://www.crypto-world.com/FactorWorld.html},~~ Webseite von Scott Contini\\
         \url{http://www.loria.fr/~zimmerma/records/factor.html}
	 \begin{sloppypar}
	     \url{http://www.emc.com/emc-plus/rsa-labs/historical/the-rsa-factoring-challenge.htm}
	 \end{sloppypar}
             % Obwohl in beiden Sprachen das usepackage url vor hyperref, wird es im Deutschen
	     % nach "-factoring-" umgebrochen; im Englischen dagegen gibt es einen schwarzen Punkt.
	     % Deshalb hier  \begin{sloppypar}    ...   \end{sloppypar} benutzt.
	     % Wenn sloppypar, dann kein // in der Anweisung zuvor.
         \url{https://www.uni-bonn.de/Pressemitteilungen/004-2010},~~ about RSA-768\\
         \url{https://members.loria.fr/PZimmermann/records/gnfs158},~~ about C158\\
         \url{http://www.loria.fr/~zimmerma/records/rsa160}

>>>>

\noindent Concretely for the RSA method does this mean:
\vskip - 1em
\begin{enumerate}
  \item It is hard to calculate $\phi(n)$ for big compounds $n$ and
  \item It is hard to calculate the prime factors of big compounds $n$
        (factorization problem).%
\footnote{%
    There is no reason for the concern sometimes mentioned that there are
    not enough primes: Raising the dimension (exponent) of the modul always
    offers enough primes to consider -- this is visualized in
    chapter~\ref{primes:_Appendix_Plotting-Primes-Quantity}
    ``\nameref{primes:_Appendix_Plotting-Primes-Quantity}''
}
\index{factorization!factorization problem}
\end{enumerate}
\vskip +1em

@@

% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\newpage
\hypertarget{SecurityRSA}{}
\section[Considerations regarding the security of the RSA algorithm]
{Regarding the security of the RSA algorithm\footnotemark}
    \footnotetext{%
    Major parts of the first part of chapter~\ref{SecurityRSA} follow
    the article ``Vorz\"uge und Grenzen des RSA-Verfahrens'' written by
    F.~Bourseau, D.~Fox, and C.~Thiel \cite{Bourseau2002}.}
    \label{SecurityRSA}

@@

%%_BEFORE_MOVING_OUT_INDEX:  \paragraph*{Factorization algorithms\footnote{%
\footnotetext{%
    \index{NT, Learning Tool for Number Theory}%
    \index{educational tool NT}%
    With the educational tool for number theory {\bf NT} you can gather more
    experience with current factorization algorithms
    (see learning unit 5.1-5.5, pages 1-15/15).\\
    NT can be called in CT1\index{CT1} via the menu path
    {\bf Indiv. Procedures \textbackslash{} Number Theory Interactive
    \textbackslash{} Learning Tool for Number Theory}.
    See appendix~\ref{s:appendix-Learn-NT}.\\
    \noindent The quadratic sieve (QS) can be found in CT1 and CT2;
    GNFS, the most modern factorization method for moduli bigger than 130
    decimal digits, is only part of CT2 (via YAFU and msieve).\\
  \noindent CT2 \index{CT2} has a GeneralFactorizer component
  based on YAFU\index{YAFU}. This is faster than the implemented functions in CT1.
  So CT2 offers the following factoring methods:
  \begin{compactitem}
   \item[-] brute-force with small primes
   \item[-] Fermat
   \item[-] Shanks square forms factorization (squfof)
   \item[-] Pollard rho
   \item[-] Pollard p-1
   \item[-] Williams p+1
   \item[-] Lenstra elliptic curve method (ECM)
   \item[-] self-initializing quadratic sieve (SIQS)
   \item[-] multiple polynomial quadratic sieve (MPQS)
   \item[-] special number field sieve (SNFS)
   \item[-] general number field sieve (GNFS).         
  \end{compactitem}
}
%%_BEFORE_MOVING_OUT_INDEX:  }
\mbox{}
%\nopagebreak

@@

An exhaustive overview about the factoring records
\index{factorization!factoring records} of composed integers using different
methods can be found on the following web pages:
\vspace{-10pt}
\begin{itemize}
\item[]
     %% URL doesn't exist any more in 2018: \url{http://www.crypto-world.com}\\
     %% http://primes.utm.edu/glossary/xpage/WheelFactorization.html no overview
     \url{http://primerecords.dk/consecutive_factorizations.htm}\\
     \url{http://en.wikipedia.org/wiki/Integer_factorization_records}\\
     \url{http://en.wikipedia.org/wiki/RSA_Factoring_Challenge}
\end{itemize}
%\footnote{%
%This site was not quite up-to-date end of May 2005: RSA-200 was missing.}

@@

The 'C numbers' originate from the Cunningham project:
\url{http://homes.cerias.purdue.edu/~ssw/cun/}.
These are factors of Mersenne numbers, which have a very special
form. This  makes it an order of magnitude easier to factor them
as moduli of the same length build for RSA.
                          }
with factorization algorithms for composed numbers are
listed in table~\ref{factorizationrecords}.

%% TODO_BE_TODO: Update 2^1193-1 laut https://www.newscientist.com/article/dn26135-factorisation-factory-smashes-number-cracking-record/  und   https://eprint.iacr.org/2014/653

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item \url{https://members.loria.fr/PZimmermann/records/gnfs158}
\item \url{https://web.archive.org/web/20170518021747/http://www.crypto-world.com:80/announcements/c158.txt}
          %% URL doesn't exist any more in 2018:
	  %% \url{http://www.crypto-world.com/announcements/c158.txt}
\end{itemize}
\end{minipage}
\vspace{12pt}

@@

On January 18, 2002 researchers at the German University of Bonn\footnote{%
          \url{https://members.loria.fr/PZimmermann/records/rsa160}\\
          \url{https://members.loria.fr/PZimmermann/records/factor.html}\\
          \url{https://web.archive.org/web/20170518021747/http://www.crypto-world.com:80/FactorWorld.html}
          %% \url{http://www.crypto-world.com/FactorWorld.html}
} 
factorized a 160-digit number into its both prime factors (these are build
with each 80 decimal digits) using the GNFS method (general number field
sieve)\index{general number field sieve (GNFS)}.

@@

\noindent\begin{minipage}{\textwidth}
\vspace{3ex}
Links:
\vspace{-10pt}
\begin{itemize}
\item \url{http://www.loria.fr/~zimmerma/records/21039-}
\item \url{https://web.archive.org/web/20170518021747/http://www.crypto-world.com:80/announcements/m1039.txt}
\item \url{https://web.archive.org/web/20170518024506/http://www.crypto-world.com:80/FactorAnnouncements.html}
\end{itemize}
\end{minipage}

@@

\noindent There are three variants:
\begin{itemize}
 \item Powerpoint 2007 (for download; dynamical, animated)\footnote{%
    \url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-en.pptx}}
 \item PDF (for download; static, no interaction)\footnote{%
    \url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-en(keine%20Interaktivitaet).pdf}}
 \item Flash (can be started within the browser, requires JavaScript; time-controlled replay)\footnote{%
\url{http://www.cryptool.org/images/ct1/presentations/RSA/RSA-Flash-en/player.html}}
\end{itemize}

@@

   \item Factorization records and factoring challenges
	 \index{factorization!factoring records}
	 \index{factorization!factoring challenges}\\
         \url{https://web.archive.org/web/20170704003418/http://www.crypto-world.com:80/},~~ last update 2013\\
         \url{https://web.archive.org/web/20170518021747/http://www.crypto-world.com:80/FactorWorld.html},~~ Webseite von Scott Contini\\
         \url{http://www.loria.fr/~zimmerma/records/factor.html}
	 \begin{sloppypar}
	     \url{http://www.emc.com/emc-plus/rsa-labs/historical/the-rsa-factoring-challenge.htm}
	 \end{sloppypar}
             % Obwohl in beiden Sprachen das usepackage url vor hyperref, wird es im Deutschen
	     % nach "-factoring-" umgebrochen; im Englischen dagegen gibt es einen schwarzen Punkt.
	     % Deshalb hier  \begin{sloppypar}    ...   \end{sloppypar} benutzt.
	     % Wenn sloppypar, dann kein // in der Anweisung zuvor.
         \url{https://www.uni-bonn.de/Pressemitteilungen/004-2010},~~ about RSA-768\\
         \url{https://members.loria.fr/PZimmermann/records/gnfs158},~~ about C158\\
         \url{http://www.loria.fr/~zimmerma/records/rsa160}
